---
title: "COVID-19 Data Analysis"
author: "M.Oshimatani"
date: "`r Sys.Date()`"
output: pdf_document
---

# Environment Setup
```{r setup, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require("ggrepel")) install.packages("ggrepel")
if(!require("sf")) install.packages("sf")
if(!require("rnaturalearth")) install.packages("rnaturalearth")
if(!require("ggthemes")) install.packages("ggthemes")
if(!require("mapproj")) install.packages("mapproj")

library(tidyverse)
library(lubridate)
library(ggrepel)
library(sf)
library(rnaturalearth)
library(ggthemes)
library(mapproj)
```

# Data Import

1. Importing Covid-19 data from the Johns Hopkins website <https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data>. Then, a vector of csv file names is passed to a variable called `file_names`, which are then combined to a single vector with `str_c()`.


```{r get_data, echo=TRUE}
url_in <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/"

file_names <- c("time_series_covid19_confirmed_global.csv", 
                "time_series_covid19_deaths_global.csv", 
                "time_series_covid19_confirmed_US.csv",
                "time_series_covid19_deaths_US.csv")

urls <- str_c(url_in, file_names)

```

2.  Read csv from each url and assigned it to a variable. 


```{r import_data, message=FALSE}
global_cases <- read_csv(urls[1])
global_deaths <- read_csv(urls[2])

```

# Data Cleansing

### Cleansing and Transforming Global Data
1. Pivot_longer to transform the cases and deaths data into one column, and then the two data frames are full joined. 

```{r global_data_clearning, message=FALSE}
global_cases <- global_cases %>% 
  pivot_longer(cols = -c(`Province/State`, `Country/Region`, Lat, Long), 
               names_to = "date", values_to = "cases") %>% 
  select(-c(Lat, Long))

global_deaths <- global_deaths %>% 
  pivot_longer(cols = -c(`Province/State`, `Country/Region`, Lat, Long), 
               names_to = "date", values_to = "deaths") %>% 
  select(-c(Lat, Long))

global <- global_cases %>% 
  full_join(global_deaths) %>% 
  rename(Country_Region = `Country/Region`, Province_State = `Province/State`)%>% 
  mutate(date = mdy(date))
```


### Add Population Data to Global Data Set
1. Create a `Combined_Key` column
```{r creating_combined_key, echo=TRUE}

global <- global %>% 
  unite("Combined_Key", 
        c(Province_State, Country_Region), 
        sep = ", ", 
        na.rm = TRUE, 
        remove = FALSE)

global
```

2. Read a population csv file from Johns Hopkins' website
```{r reading_poplation, message=FALSE}
uid_lookup_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/UID_ISO_FIPS_LookUp_Table.csv"

uid <- read_csv(uid_lookup_url) %>% 
  select(-c(Lat, Long_, Combined_Key, code3, iso2, iso3, Admin2))

# Checking for NA in population
uid[is.na(uid$Population),]

```

4. Selecting the top row population for each country. This is important because it lists the total population in the first row, followed by the breakdowns of its provinces. 

```{r cleaning_up_pop_data}
uid_cleaned <- uid %>%
  group_by(Country_Region) %>%
  summarise(Population = first(Population))

uid_cleaned
```
3. Add the population data into global data
```{r adding_pop_data, echo=TRUE}
global_pop <- global %>% 
  left_join(uid_cleaned, by="Country_Region") 

global_pop
```

4. Create `global_total` df

```{r global_by_country, message=FALSE}
global_total <- global_pop %>%
  group_by(Country_Region, date) %>%
  summarize(cases = sum(cases), 
            deaths = sum(deaths), 
            Population = sum(Population)) %>%
  mutate(deaths_per_mill = deaths * 1000000 / Population,
         cases_per_mill = cases * 1000000 / Population) %>% 
  select(Country_Region, date, cases, deaths, 
         deaths_per_mill, cases_per_mill, Population) %>% 
  ungroup()
  
global_total
```

5. Handle NA
```{r na_handling, message=FALSE}

complete_rows <- complete.cases(global_total)
global_complete <- global_total[complete_rows,]

summary(global_complete)

```

# Visualization 1

1. Filter `global_complete` to extract Japan data.

```{r japan_data, warning=FALSE}
japan <- global_complete %>% filter(Country_Region == "Japan")

  
```

2. Visualize Covid19 in Japan
```{r japan_visualization, warning=FALSE}
ggplot(data=japan, aes(x=date, y=deaths)) + 
  geom_line(aes(colour = "deaths")) + 
  geom_point(aes(color="deaths"))+
  geom_line(aes(y=cases, colour="cases")) +
  geom_point(aes(y=cases, colour="cases")) +
  scale_y_log10() +
  theme(legend.position="bottom")+
  labs(title = "COVID19 in Japan", x = "Date", y=NULL)
```

3. Calculate Cases per million, Deaths per million, mortality rate of Japan
```{r japan_stats, message=FALSE}
japan_max <-japan %>% top_n(1, date)
cat("Cases per million for Japan is ", japan_max$cases_per_mill, "\n\n")
cat("Deaths per million for Japan is ", japan_max$deaths_per_mill, "\n\n")
cat("Mortality Rate in Japan is ",  japan_max$deaths / japan_max$cases,  "\n\n")
```
Cases per million for Japan is 263451 and deaths per million is 557 as of 03-09-2023. 

# Visualization 2

1. Extract worst 10 countries in deaths per million

```{r global_worst10, message=FALSE}
worst10_data <- global_complete %>% 
  filter(date == "2023-03-09") %>%
  slice_max(deaths_per_mill, n=10)
worst10 <- worst10_data$Country_Region
cat("Worst 10 countries are: ", worst10)

```

2. Visualize worst 10 countries

```{r worst10_viz, warning=FALSE}
worst10_df <- global_complete %>%
  filter(Country_Region %in% worst10)

worst10_data <- worst10_data %>% mutate(deaths_per_mill = round(deaths_per_mill))
ggplot(data=worst10_df, aes(x=date, y=deaths_per_mill, group = Country_Region))+
  geom_line(aes(colour = Country_Region)) +
  geom_text_repel(
    aes(label = deaths_per_mill), data = worst10_data ,
    fontface ="plain", color = "black", size = 3
    )+
  labs(title = "Worst 10 Countries", x = "Date", y="Deaths Per Million")
```

# Map Visualization

1. Summarize the data
```{r transforming_data, message=FALSE}
global_country_totals <- global_complete %>% 
  group_by(Country_Region) %>% 
  summarize(deaths = max(deaths), 
            cases = max(cases), 
            population=max(Population), 
            cases_per_thou = 1000 * cases/population, 
            deaths_per_thou = 1000*deaths/population) %>% 
  filter(cases > 0, population > 0)

global_country_totals
```


2. Download world map data and resolve the names between the `region` of the map and `Country_Region` of the global_deaths df. 
```{r dowload_world_data}
world_map <- map_data("world")%>%
  filter(! long > 180)

countries <- world_map %>%
  distinct(region) %>%
  rowid_to_column()

global_deaths <- global_country_totals %>%
  select(Country_Region, deaths_per_thou) %>%
  mutate(Country_Region = recode(str_trim(Country_Region),
                                 "US" = "USA",
                                 "United Kingdom"="UK",
                                 "Korea, South" = "South Korea",
                                 "Congo (Brazzaville)"  ="Republic of Congo",
                                 "Congo (Kinshasa)" = "Democratic Republic of the Congo",
                                 "Korea, North" = "North Korea",
                                 "Czechia" = "Czech Republic"))
global_deaths

```


3. Visualize the map
```{r map, warning=FALSE}

global_deaths %>%
  ggplot(aes(fill = deaths_per_thou, map_id = Country_Region)) +
  geom_map(map = world_map) +
  expand_limits(x = world_map$long, y = world_map$lat) +
  scale_fill_continuous(high = "#132B43", low = "#56B1F7")+
  coord_map("mercator") +
  theme_map()+
  labs(title="Global Map of Covid-19 Deaths Per Thousand", fill="Deaths per thousand")
```

You can see the worst 10 countries in the Visualization 2 are darker color on this world map. 

# Modeling the Data
1. Create a linear model

```{r linear_model, echo=TRUE}
mod <- lm(deaths_per_thou ~ cases_per_thou, data=global_country_totals)
summary(mod)

```

2. Analyze the data

-- Which country has the smallest cases per thousand?

```{r smallest_cpt}
global_country_totals %>% slice_min(cases_per_thou)
```


 -- Which country has the largest cases per thousand?
```{r largest_cpt}
global_country_totals %>% slice_max(cases_per_thou)
```
3. Add prediction to the data and visualize it.

```{r ggplot_pred, include=TRUE}
global_tot_w_pred <- global_country_totals %>% mutate(pred = predict(mod))

global_tot_w_pred %>% ggplot()+
  geom_point(aes(x = cases_per_thou, y=deaths_per_thou), color="blue") + 
  geom_point(aes(x=cases_per_thou, y=pred), color="red") +
  labs(title = "Global COVID-19 Linear Model", x="Cases per thousand", y="Death per thousand")

```

# Summary
In this analysis, I focused on global COVID-19 data. First, I identified the 10 countries with the highest death rates per million and visualized their trends. Second, I mapped death rates per million globally, highlighting the top 10 countries with darker colors.

To model the relationship between cases and deaths, I used a linear model. The model's adjusted R-squared was 0.277, so it was relatively weak fit.  But still, there is a clear correlation between the variables.

As for biases, there's some biases involved in this analysis.  First one is outlier handling bias.  I excluded Antarctica because of two reasons.  The primary reason is that its population data was missing.  This was crucial for calculating deaths per million and cases per millions for further study.  And another reason is that, it's a unique environment with small population mostly composed of researchers. This makes it an outlier compared to other regions. 

As a next step, I recommend a deeper dive into the data from one of the top 10 countries to explore further trend and patterns. 

